Draft

### 2017-09-04 Thesis/Draft.Rmd
Edit second chapter

@@

# Information Theoretic Complexity of Pure Exploration Bandits \label{chap:KullbackLeiblerPureExploration}

@@

## Complexity Measures for Best-Arm and Top-*m* Identification \label{sec:KaufmannEtAl2016}

Instead of the thresholding bandit problem, Kaufmann et al. (2016) consider the problem of identifying the best $m$ out of $K$ arms, that is, the Top-$m$ problem. They describe their a bandit model $\nu$ also by $K$ arms characterized by probability distributions $\nu_a$, $a \in \{1,...,K\}$ with mean $\mu_a$. As before, a learner decides at any time step $t=1,2,...,$ an arm $A_t \in \{1,...,K\}$ for which he observes the independent reward $Z_{t} \sim \nu_{A_t}$. The learner aims to find the $m$ out of $K$ arms with the largest expected reward. This set of arm indices is denoted by $\mathcal{S}_m^*$. Kaufmann et al. write the in decreasing order sorted means $(\mu_{1}, ..., \mu_{K})$ as the K-tuple $(\mu_{[1]}, ..., \mu_{[K]})$. Based on this, they require that $\mu_{[m]} > \mu_{[m+1]}$ such that the solution $\mathcal{S}_m^*$ is unique. The class of bandit problems $\nu$ for which the latter condition holds is called $\mathcal{M}_m$ by Kaufmann et al. At the end of the sampling procedure, after $T$ samples (denoted by $\tau$ in Kaufmann et al.) the learner is required to return a classification $\hat{\mathcal{S}}_m$. This classification is then evaluated by the loss function $\mathcal{L}(T) = \mathbb{I}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m)$. This binary loss is of course very similar to the loss in the thresholding bandit problem. We can now define a strategy of the learner as $\mathcal{A} = ((A_t), T, \hat{\mathcal{S}}_m)$: The strategy defines how he chooses an arm, how many samples are observed, and how the arms are classified in the end.

When introducing the thresholding bandit setting, we did not mention the fact that the budget $T$ does not need to be fixed upfront. Indeed, the setting in which the budget is *not* fixed upfront might be more popular among the best arm identification and top-$m$ literature. This setting is called *fixed confidence*, while the setting we considered so far is called *fixed budget*. In the fixed budget setting, we run algorithms for $T$ samples and then compare them based on their expected loss, $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m) =: p_t(\nu)$ as we have seen in $\autoref{sec:Setup}$. Alternatively, consider a $\delta$-PAC strategy $\mathcal{A}(\delta)$for which holds for every problem $\nu \in \mathcal{M}_m$ that the classification holds at least with confidence level $1-\delta$, that is, $\mathbb{P}(\mathcal{S}^*_m = \hat{\mathcal{S}}_m) \geq 1-\delta$. The strategy has to achieve this level of confidence in its prediction before it is allowed to stop sampling. This is the fixed confidence case, in which the budget $T$ is no longer fixed upfront, but becomes a random variable. We compare algorithms by their expected number of draws $\mathbb{E}_{\nu}[T_{\delta}]$ and prefer those which arrive at the confidence level $1-\delta$ with a smaller expected budget.

Following Kaufmann et al. (2016), in the fixed confidence setting, a family of strategies $A = (\mathcal{A}(\delta))_{\delta \in (0,1)}$ is called PAC if for every $\delta$, $\mathcal{A}(\delta)$ is $\delta$-PAC. On the other hand, in the fixed budget setting, a family of strategies $A = (\mathcal{A}(t))_{t \in \mathbb{N}}$ is called consistent if $p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$ for every problem $\nu \in \mathcal{M}_m$.

Given this setup, we can now follow the arguments of Kaufmann et al. (2016) to deduce complexity measures for the top-$m$ problem. As alluded to earlier, these complexity measures will rely on the Kullback-Leibler divergence in contrast to the complexity $H$ used by Locatelli et al. (2016). Also, when discussing $H$, we so far described it as complexity as it measures the difficulty of the problem by measuring the distance between arms and the threshold. Kaufmann et al. approach complexity instead from how many samples (what error probability) one can expect to need in the fixed confidence (fixed budget) setting. Thus, in order to arrive at their new measure of complexity, it makes sense to first follow their arguments to a general lower bound on the sample complexity $\mathbb{E}_{\nu}[\tau]$ in the case of fixed confidence top-$m$ problems. This bound is given in their Theorem 4 (for common confidence levels of $\delta \leq 0.15$).

@@

\begin{equation*}
\mathbb{E}_{\nu}[\tau] \geq \Big[ \sum_{a \in \mathcal{S}^*_m} \frac{1}{KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}^*_m} \frac{1}{KL(\nu_a, \nu_{[m]})} \Big] \der(1-\delta, \delta),
\end{equation*}

where $\der(x,y) := x \log(x/y) + (1-x) \log((1-x)/(1-y))$ is the binary relative entropy, with the convention that $\der(0,0) = \der(1,1) = 0$.
\end{theorem}

To prove this statement for a given $\delta$-PAC algorithm $\mathcal{A}$, we assume certain continuity conditions (compare Kaufmann et al. (2016)), with which there exists for every arm $a \in \{1, \dots, K\}$ an alternative model $\nu' = (\nu_1, \dots, \nu_{a-1}, \nu_a', \nu_{a+1}, \dots, \nu_K)$. In this alternative model, in accordance with the continuity conditions, only arm $a$ is modified. After the modification, the arm $\nu_{m+1}$ is now closer to $\nu_a$ as measured by the Kullback-Leibler divergence than $\nu_a'$ is to $\nu_a$, if $\mu_a > \mu_{m+1}$ (consequently, $\mu_a' < \mu_{m+1}$). On the other hand, if $\mu_a < \mu_{m}$, then $\nu_a'$ will be moved away from $\nu_a$ just enough so that $\nu_a$ is closer to $\nu_m$ than to $\nu_a'$ (consequently, $\mu_a' > \mu_m$). These changes in distribution make the original set of optimal arms, $\{1,\dots,m\}$, no longer optimal under the alternative model $\nu'$. To compare this to the thresholding bandit problem, assume that arms $\nu_{m}$ and $\nu_{m+1}$ indirectly define a threshold $\tau \in (\mu_{m+1},\mu_m)$. Then, by creating an alternative problem as described here, we flip arm $a$ around the implied threshold $\tau$ (compare the lower bound proof in \autoref{sec:AppendixAPTLB}).

Given that the strategy returns $\hat{S}_m$, we know that for the event $\mathcal{E} = (\hat{S}_m = \{1, \dots, m\})$ (which is element of the $\sigma$-Algebra $\mathcal{F}_{T}$ induced by the stopping time $T$), that $\mathbb{P}_{\nu}(\mathcal{E}) \geq 1-\delta$ and $\mathbb{P}_{\nu'}(\mathcal{E}) \leq \delta$ (because fixed confidence $\delta$!). To now give the lower bound on the expected number of samples, $\mathbb{E}_{\nu}[\tau]$, we employ the general Lemma given by the authors for such changes in distribution in bandit models.

\begin{lemma}[Kaufmann et al., 2016] \label{theorem:KaufmannEtAlLemma1}
Let $\nu$ and $\nu'$ be two bandit models with $K$ arms such that for all $a$, the distributions $\nu_a$ and $\nu_a'$ are mutually absolutely continuous. For any almost-surely finite stopping time $\sigma$ with respect to $(\mathcal{F}_t)$,

\begin{equation*}
\sum_{a=1}^{K} \mathbb{E}_{\nu} [N_a(\sigma)] \KL(\nu_a, \nu_a') \geq \sup_{\mathcal{E} \in \mathcal{F}_{\sigma}} \der (\mathbb{P}_{\nu}(\mathcal{E}), \mathbb{P}_{\nu'}(\mathcal{E})),
\end{equation*}

where $d(x,y)$ is the binary relative entropy as defined above.
\end{lemma}

We can apply this Lemma directly on the expected number of draws for each arm, $\mathbb{E}_{\nu}[N_a]$. For each arm $a$, we thus have $\KL(\nu_a, \nu_a') \mathbb{E}_{\nu}[N_a] \geq d(1-\delta, \delta)$. Using the definition of the alternative model above, we get for every small but fixed $\alpha > 0$, for every arm $a \in \{1, \dots, K\}$ and each arm $b \in \{m+1, \dots, K\}$:

\begin{equation*}
\mathbb{E}_{\nu}[N_a] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_a, \nu_{m+1}) + \alpha}
\end{equation*}

and

\begin{equation*}
\mathbb{E}_{\nu}[N_b] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_b, \nu_{m}) + \alpha}.
\end{equation*}

We can then let $\alpha \to 0$, sum the bounds on the individual arms to get the bound on the overall number of draws:

\begin{equation*}
\mathbb{E}_{\nu}[T] = \sum_{a=1}^K \mathbb{E}_{\nu}[N_a] \geq \Big( \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})} \Big) \der(1-\delta, \delta)
\end{equation*}

@@

\begin{equation*}
\mathbb{E}_{\nu}[T] \geq \Big( \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})} \Big) \log \big(\frac{1}{2.4\delta} \big).
\end{equation*}

Given this lower bound, it becomes clear that the complexity of the top-$m$ problem can be lower bounded by the Kullback-Leibler divergence between arms and the implied threshold between the $m$-th best arm and the $(m+1)$-th best arm. After all, the latter gives a lower bound on the effort (in terms of expected number of samples) required to achieve a certain confidence level. And a problem that requires a larger effort can be described as more complex. Indeed, in the case of fixed confidence problems, Kaufmann et al. (2016) describe problems as having a larger *sampling complexity*.

Following Kaufmann et al. (2016), we can now define the complexity $\kappa_C(\nu)$ for the fixed confidence setting as

\begin{equation}
\kappa_C(\nu) = \inf_{A \, \text{PAC}} \lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\nu}[T{\delta}]}{\log \frac{1}{\delta}}. \label{eq:KappaC}
\end{equation}

@@

\begin{equation*}
\kappa_C(\nu) \geq \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})}
\end{equation*}

Given this description, we would expect that a complexity measure in a fixed confidence thresholding bandit probem might involve replacing $\mu_{[m]}$ and $\mu_{[m+1]}$ by the actual threshold $\tau$ that divides the group above the threshold from the group below the threshold.

However, given the problem formulation in Locatelli et al. (2016), we are mainly interested in complexity measures and algorithms for the fixed budget setting. As we will see now, the results of Kaufmann et al. (2016) are more limited for this case. While the authors are able to provide a bound on their complexity measure in the fixed budget case, it only applies to the case of *two-armed* bandit models. Consider a consistent fixed budget strategy $\mathcal{A}$ ($p_t(\nu)$ tends to zero as $t \rightarrow \infty$), and define the complexity in the fixed budget setting as follows:

\begin{equation}
\kappa_B(\nu) = \inf_{A \, \text{consistent}} \big(\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu)\big)^{-1}.
\end{equation}

While we included the experimenter-fixed $\delta$ in the definition of $\kappa_C$, we now include the experimenter-fixed $T$ in the definition of $\kappa_B$ to disregard it in the complexity measure after plugging in the lower bound on $p_t(\nu)$. The resulting lower bound given in Theorem 12 of Kaufmann et al. (2016):

@@

\begin{equation*}
\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu) \leq \inf_{(\nu_1', \nu_2') \in \mathcal{M}: \mu_1' < \mu_2'} \frac{\KL(\nu_1', \nu_1) + \KL(\nu_2', \nu_2)}{2}
\end{equation*}
\end{theorem}

For a proof, consider first the two different bandit models, $\nu$ and $\nu'$, where we fix w.l.o.g. $a^* = 1$ as the best arm in $\nu$, while $a^* = 2$ in $\nu'$. Note that in this case we don't necessarily try to fool the algorithm by a change of distribution. Indeed, we want the algorithm $\mathcal{A}$ to be consistent for both problems. Since we are in the fixed budget setting, the stopping time is fixed at $T = t$. Even though the correct decision differs between the two problems, we focus on the event $A = (\hat{S}_1 = 1)$. Now, we can apply Lemma 1 as before, as it in essence gives a measure of how different two bandit models are. Here, we want to know how different $\nu$ is from $\nu'$, and so, given we have only two arms,

\begin{equation*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2', \nu_2) \geq \der(\mathbb{P}_{\nu'}(A),\mathbb{P}_{\nu}(A)))
\end{equation*}

@@

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2', \nu_2) & \geq \der(\epsilon,1-p_t(\nu))) \\
& = \epsilon \log \frac{\epsilon}{1-p_t(\nu)} + (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} \\
& \geq (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} + \epsilon \log \epsilon.
\end{align*}

@@

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2', \nu_2)
& \geq \log \frac{1}{p_t(\nu)} = -\log p_t(\nu).
\end{align*}

@@

\begin{align*}
\lim_{t \to 0} \sup - \frac{1}{t} \log p_t(\nu)
& \leq \lim_{t \to 0} \sup \frac{1}{t} \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)] \KL(\nu_a', \nu_a) \\
& \leq \max_{a=1,2} \KL(\nu_a', \nu_a),
\end{align*}

@@

Note that while deriving the lower bound on the complexity measure, we have also derived the following lower bound on the failure probability for the two-armed fixed budget bandit:

\begin{equation*}
p_t(\nu) \geq \exp \big(- \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)] \KL(\nu_a', \nu_a) \big)
\end{equation*}

or, to not rely on the share of samples attributed to each arm,

\begin{equation*}
p_t(\nu) \geq \exp \big(- \max_{a=1,2} \KL(\nu_a', \nu_a) \big).
\end{equation*}

@@

Before we discuss the strategies of Kaufmann et al. (2016) matching the lower bounds, we now show how lower bounds can be derived very similarly for the fixed budget thresholding problem.

### Lower Bound for Fixed Budget Thresholding Problem

In order to derive a bound for the fixed budget problem, Kaufmann et al. (2016) have focused on the case of the two-armed bandit which is the most simple version of the top-$m$ problem. Less than two arms do not make sense, since then this single arm will always be the best arm. In the case of the thresholding bandit, however, a single-armed problem makes still some sense in that we still have to classify this arm correctly with respect to the threshold. Thus, we can show the following corollary which gives the a bound on the complexity of the single-armed thresholding bandit.

\begin{corollary}
Let $\nu = \nu_1$ be a one-armed thresholding bandit model where $\tau$ is the threshold fixed upfront. Let $\nu$ be from the univariate canonical exponential family of distributions such that it can be parameterized by its mean, $\mu_1$, and the Kullback-Leibler divergence between distributions $\nu_1$ and $\nu_1'$ can be written as $\KL(\nu_1', \nu_1) = \KL(\mu_1', \mu_1)$. In the fixed budget setting, any consistent algorithm satisfies:

$$
- \frac{1}{t} \log p_t(\nu) \leq \KL(\tau, \mu_1)
$$
\end{corollary}

This of course is basically the Chernoff-Stein Lemma for the hypothesis test of two alternatives (compare chapter 11 in Cover and Thomas, 2006).

But as we see in the following corollary, we can extend the lower bound to the $K$-armed fixed budget thresholding bandit problem, which stands in contrast to the fact that for the top-$m$ problem, it is only possible to show the lower bound for two arms.

\begin{corollary}
Let $\nu = (\nu_1, ..., \nu_K)$ be a $K$-armed thresholding bandit problem in which $\tau$ is the threshold fixed upfront. For $i \in \{1,...,K\}$, let $\nu_i$ be from the univariate canonical exponential family of distributions such that it can be parameterized by its mean, $\mu_i$, and the Kullback-Leibler divergence between distributions $\nu_i$ and $\nu_i'$ can be written as $\KL(\nu_i, \nu_i') = \KL(\mu_i, \mu_i')$. In the fixed budget setting, every consistent algorithm satisfies:

$$
\lim_{t \rightarrow \infty} -\frac{1}{t} \log p_t(\nu) \leq \max_{a \in \{1,...,K\}} \KL(\tau, \mu_a)
$$
\end{corollary}

Given the lower bounds on the sampling complexity and error probability, as well as the complexity measures, we now turn to sampling strategies that are designed to match these lower bounds.

### Matching algorithm strategies

>>>>

Define exponential family of distributions

Define KL Divergence for this family

@@

# Information Theoretic Approaches to Pure Exploration Bandits \label{chap:KullbackLeiblerPureExploration}

@@

## Information Theoretic Complexity Measures in Top-$m$ Problems \label{sec:KaufmannEtAl2016}

Instead of the thresholding bandit problem, Kaufmann et al. (2016) consider the problem of identifying the best $m$ out of $K$ arms, that is, the top-$m$ problem. They describe their bandit model $\nu$ by $K$ arms characterized by probability distributions $\nu_a$, $a \in \{1,...,K\}$, with mean $\mu_a$. A learner picks at any time step $t=1,2,...,$ an arm $A_t \in \{1,...,K\}$ for which he observes the independent reward $Z_{t} \sim \nu_{A_t}$. The learner aims to find the $m$ out of $K$ arms with the largest expected reward. This set of arm indices is denoted by $\mathcal{S}_m^*$. Kaufmann et al. write the means $(\mu_{1}, ..., \mu_{K})$ sorted in decreasing order as the K-tuple $(\mu_{[1]}, ..., \mu_{[K]})$. Based on this, they require that $\mu_{[m]} > \mu_{[m+1]}$ such that the solution $\mathcal{S}_m^*$ is unique. The class of bandit problems $\nu$ for which the latter condition holds is called $\mathcal{M}_m$ by Kaufmann et al. At the end of the sampling procedure, after $T$ samples (denoted by $\tau$ in Kaufmann et al.) the learner is required to return a classification $\hat{\mathcal{S}}_m$. This classification is then evaluated by the loss function $\mathcal{L}(T) = \mathbb{I}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m)$. This binary loss is similar to the loss in the thresholding bandit. We can now define a strategy of the learner as $\mathcal{A} = ((A_t), T, \hat{\mathcal{S}}_m)$: The strategy defines how an arm is chosen, how many samples are observed, and how the arms are eventually classified.

When introducing the thresholding bandit setting, we did not mention the fact that the budget $T$ does not need to be fixed upfront. Indeed, the setting in which the budget is *not* fixed upfront might be more popular among the best arm identification and top-$m$ literature. This setting is called *fixed confidence*, while the setting we considered so far is called *fixed budget*. In the fixed budget setting, we run algorithms for $T$ samples and then compare them based on their expected loss, $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m) =: p_t(\nu)$ as we have seen in $\autoref{sec:Setup}$. Alternatively, consider a $\delta$-PAC strategy $\mathcal{A}(\delta)$for which it holds for every problem $\nu \in \mathcal{M}_m$ that the classification holds at least with confidence level $1-\delta$, that is, $\mathbb{P}(\mathcal{S}^*_m = \hat{\mathcal{S}}_m) \geq 1-\delta$. The strategy has to achieve this level of confidence in its prediction before it is allowed to stop sampling. This is the fixed confidence case, in which the budget $T$ is no longer fixed upfront, but turns into a random variable. Algorithms are then compared by their expected number of draws $\mathbb{E}_{\nu}[T_{\delta}]$ and we prefer those which achieve the confidence level $1-\delta$ with a smaller expected budget.

In the fixed confidence setting, a family of strategies $A = (\mathcal{A}(\delta))_{\delta \in (0,1)}$ is called PAC if for every $\delta$, $\mathcal{A}(\delta)$ is $\delta$-PAC. On the other hand, in the fixed budget setting, a family of strategies $A = (\mathcal{A}(t))_{t \in \mathbb{N}}$ is called consistent if $p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$ for every problem $\nu \in \mathcal{M}_m$.

Given this setup, we can now follow the arguments of Kaufmann et al. (2016) to derive complexity measures for the top-$m$ problem. As alluded to earlier, these complexity measures will rely on the KL-divergence in contrast to the complexity $H$ used by Locatelli et al. (2016). When discussing $H$, its motivation is that it represents the difficulty of the problem by measuring the distance between arms and the threshold. Instead, Kaufmann et al. approach complexity by how many samples (what error probability) one will need in the fixed confidence (fixed budget) setting. In order to arrive at their new measure of complexity, it makes sense to first follow their arguments to a general lower bound on the sample complexity $\mathbb{E}_{\nu}[\tau]$ which is given in their Theorem 4:

@@

\begin{equation*}
\mathbb{E}_{\nu}[\tau] \geq \Big[ \sum_{a \in \mathcal{S}^*_m} \frac{1}{\KL(\nu_a || \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}^*_m} \frac{1}{\KL(\nu_a || \nu_{[m]})} \Big] \der(1-\delta, \delta),
\end{equation*}
\end{theorem}

To prove this statement for a given $\delta$-PAC algorithm $\mathcal{A}$, we assume certain continuity conditions (compare Kaufmann et al. (2016)) under which there exists for every arm $a \in \{1, \dots, K\}$ an alternative model $\nu' = (\nu_1, \dots, \nu_{a-1}, \nu_a', \nu_{a+1}, \dots, \nu_K)$. In this alternative model, in accordance with the continuity conditions, only arm $a$ is modified. After the modification, the arm $\nu_{m+1}$ is now closer to $\nu_a$ as measured by the Kullback-Leibler divergence than $\nu_a'$ is to $\nu_a$, if $\mu_a > \mu_{m+1}$ (consequently, $\mu_a' < \mu_{m+1}$). On the other hand, if $\mu_a < \mu_{m}$, then $\nu_a'$ will be moved away from $\nu_a$ just enough so that $\nu_a$ is closer to $\nu_m$ than to $\nu_a'$ (consequently, $\mu_a' > \mu_m$). These changes in distribution make the original set of optimal arms, $\{1,\dots,m\}$, no longer optimal under the alternative model $\nu'$. To compare this to the thresholding bandit problem, assume that arms $\nu_{m}$ and $\nu_{m+1}$ indirectly define a threshold $\tau \in (\mu_{m+1},\mu_m)$. Then, by creating an alternative problem as described here, we flip arm $a$ around the implied threshold $\tau$ (compare the lower bound proof in \autoref{sec:AppendixAPTLB}).

Given the strategy returns $\hat{S}_m$, we know that for the event $\mathcal{E} = (\hat{S}_m = \{1, \dots, m\})$ (which is element of the $\sigma$-Algebra $\mathcal{F}_{T}$ induced by the stopping time $T$), that $\mathbb{P}_{\nu}(\mathcal{E}) \geq 1-\delta$ and $\mathbb{P}_{\nu'}(\mathcal{E}) \leq \delta$ (because fixed confidence $\delta$!). To give the lower bound on the expected number of samples, $\mathbb{E}_{\nu}[\tau]$, we employ the general Lemma given by the authors for such changes in distribution in bandit models.

\begin{lemma}[Lemma 1, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlLemma1}
Let $\nu$ and $\nu'$ be two bandit models with $K$ arms such that for all $a$, the distributions $\nu_a$ and $\nu_a'$ are mutually absolutely continuous. For any almost-surely finite stopping time $\sigma$ with respect to $(\mathcal{F}_t)$,

\begin{equation*}
\sum_{a=1}^{K} \mathbb{E}_{\nu} [N_a(\sigma)] \KL(\nu_a|| \nu_a') \geq \sup_{\mathcal{E} \in \mathcal{F}_{\sigma}} \der (\mathbb{P}_{\nu}(\mathcal{E}), \mathbb{P}_{\nu'}(\mathcal{E})).
\end{equation*}
\end{lemma}

We apply this Lemma directly on the expected number of draws for each arm, $\mathbb{E}_{\nu}[N_a]$. For each arm $a$, we thus have $\KL(\nu_a, \nu_a') \mathbb{E}_{\nu}[N_a] \geq d(1-\delta, \delta)$. Using the definition of the alternative model above, we get for every small but fixed $\alpha > 0$, every arm $a \in \{1, \dots, K\}$ and $b \in \{m+1, \dots, K\}$:

\begin{equation*}
\mathbb{E}_{\nu}[N_a] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_a || \nu_{m+1}) + \alpha} \quad \text{and} \quad \mathbb{E}_{\nu}[N_b] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_b || \nu_{m}) + \alpha}.
\end{equation*}

Let $\alpha \to 0$, sum the individual arms' bounds to get the bound on the overall number of draws:

\begin{equation*}
\mathbb{E}_{\nu}[T] = \sum_{a=1}^K \mathbb{E}_{\nu}[N_a] \geq \Big( \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a|| \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a|| \nu_{[m]})} \Big) \der(1-\delta, \delta)
\end{equation*}

@@

\begin{equation*}
\mathbb{E}_{\nu}[T] \geq \Big( \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a|| \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a|| \nu_{[m]})} \Big) \log \big(\frac{1}{2.4\delta} \big).
\end{equation*}

Given this lower bound, it is clear that the complexity of the top-$m$ problem can be lower bounded by the KL-divergence between arms and the implied threshold between the $m$-th best arm and the $(m+1)$-th best arm. The latter gives a lower bound on the effort (in terms of expected number of samples) required to achieve a certain confidence level, where a problem that requires a larger effort is described as more complex. Indeed, in the case of fixed confidence problems, Kaufmann et al. (2016) describe problems as having a larger *sampling complexity*.

Following Kaufmann et al. (2016), we define the complexity $\kappa_C(\nu)$ in the fixed confidence setting as

\begin{equation}
\kappa_C(\nu) = \inf_{A \, \text{PAC}} \limsup_{\delta \to 0} \frac{\mathbb{E}_{\nu}[T_{\delta}]}{\log \frac{1}{\delta}}. \label{eq:KappaC}
\end{equation}

@@

\begin{equation*}
\kappa_C(\nu) \geq \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a || \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a || \nu_{[m]})}
\end{equation*}

Given this description, we would expect that a complexity measure in a fixed confidence thresholding bandit probem might involve replacing $\mu_{[m]}$ and $\mu_{[m+1]}$ by the actual threshold $\tau$ that divides the group above the threshold from the group below the threshold. But given the setting in Locatelli et al. (2016), we are mainly interested in complexity measures and algorithms for the fixed budget problem. The results of Kaufmann et al. (2016) are more limited for this case. While the authors are able to provide a bound on their complexity measure in the fixed budget case, it only applies to the case of *two-armed* bandit models.

Consider a consistent fixed budget strategy $\mathcal{A}$ ($p_t(\nu)$ tends to zero as $t \rightarrow \infty$), and define the complexity in the fixed budget setting as follows:

\begin{equation}
\kappa_B(\nu) = \inf_{A \, \text{consistent}} \Big(\limsup_{t \to \infty} - \frac{1}{t} \log p_t(\nu)\Big)^{-1}.
\end{equation}

While we included the experimenter-fixed $\delta$ in the definition of $\kappa_C$, we now include the fixed $T$ in the definition of $\kappa_B$ to disregard it in the complexity measure after plugging in the lower bound on $p_t(\nu)$. The resulting lower bound given in Theorem 12 of Kaufmann et al. (2016):

@@

\begin{equation*}
\limsup_{t \to \infty} - \frac{1}{t} \log p_t(\nu) \leq \inf_{(\nu_1', \nu_2') \in \mathcal{M}: \mu_1' < \mu_2'} \max \big\{\KL(\nu_1' || \nu_1), \KL(\nu_2' || \nu_2)\big\}
\end{equation*}
\end{theorem}

For a proof, consider first the two different bandit models, $\nu$ and $\nu'$, where we fix w.l.o.g. $a^* = 1$ as the best arm in $\nu$, while $a^* = 2$ in $\nu'$. We require the algorithm $\mathcal{A}$ to be consistent for both problems. Since we are in the fixed budget setting, the stopping time is fixed at $T = t$. The correct decision differs between the two problems, and we focus on the event $A = (\hat{S}_1 = 1)$. Now, we can apply Lemma 1 from Kaufmann et al. (2016) as before, as it gives a measure of how different two bandit models are (it bounds the change in distribution). Here, we want to know how different $\nu$ is from $\nu'$, and, given we have only two arms,

\begin{equation*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1'|| \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2'|| \nu_2) \geq \der(\mathbb{P}_{\nu'}(A),\mathbb{P}_{\nu}(A)))
\end{equation*}

@@

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1' || \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2' || \nu_2) & \geq \der(\epsilon,1-p_t(\nu))) \\
& = \epsilon \log \frac{\epsilon}{1-p_t(\nu)} + (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} \\
& \geq (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} + \epsilon \log \epsilon.
\end{align*}

@@

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1' || \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2' || \nu_2)
& \geq \log \frac{1}{p_t(\nu)} = -\log p_t(\nu).
\end{align*}

@@

\begin{align*}
\limsup_{t \to \infty} - \frac{1}{t} \log p_t(\nu)
& \leq \limsup_{t \to 0} \frac{1}{t} \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)] \KL(\nu_a' || \nu_a) \\
& \leq \max_{a=1,2} \KL(\nu_a' || \nu_a),
\end{align*}

@@

@@

Before we discuss algorithms given in Kaufmann et al. (2016) that match the lower bounds, we show how lower bounds can be derived very similarly for the fixed budget thresholding problem.

### Lower Bound for Fixed Budget Thresholding Problem \label{sec:LBTBP}

In order to derive a bound for the fixed budget problem, Kaufmann et al. (2016) have focused on the setting with two arms which is the most simple version of the top-$m$ problem. Less than two arms do not make sense, since the single arm is by definition the best arm. In the case of the thresholding bandit, however, a single-armed problem makes still at least limited sense in that we still have to classify this arm correctly with respect to the threshold. Thus, we can show the following corollary which gives a bound on the complexity of the single-armed thresholding bandit.

\begin{corollary}
Let $\nu = \nu_1$ be a one-armed thresholding bandit model where $\tau$ is the threshold fixed upfront. Let $\nu$ be from the univariate exponential family of distributions such that $\nu_1$ can be parameterized by its mean, $\mu_1$, and the Kullback-Leibler divergence between distributions $\nu_1$ and $\nu_1'$ can be written as $\KL(\nu_1', \nu_1) = \KL(\mu_1', \mu_1)$. In the fixed budget setting, any consistent algorithm ($p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$) satisfies:

$$
- \frac{1}{t} \log p_t(\nu) \leq \KL(\tau || \mu_1)
$$
\end{corollary}\label{corollary:LB1TBP}

This is basically the Chernoff-Stein Lemma for the hypothesis test of two alternatives (compare chapter 11 in Cover and Thomas, 2006, or \autoref{theorem:ChernoffStein}). We arrive at this result as the one-armed bandit does not require sequential sampling: The single arm receives all samples. $\mathbb{E}_{\nu'}[N_1(t)]$ is not a random variable but equals $T$.

But as we see in the following corollary, we can extend the lower bound to the $K$-armed fixed budget thresholding bandit problem, which stands in contrast to the fact that for the top-$m$ problem, it is only possible to show the lower bound for two arms. The result is based on an alternative setting $\nu'$ in which all arms have means $\mu' = \tau$ such that their expected number of pulls corresponds to uniform sampling, $\mathbb{E}_{\nu'}[N_i(t)] = T/K$.

\begin{corollary}
Let $\nu = (\nu_1, ..., \nu_K)$ be a $K$-armed thresholding bandit problem in which $\tau$ is the threshold fixed upfront. For $i \in \{1,...,K\}$, let $\nu_i$ be from the univariate exponential family of distributions such that $\nu_i$ can be parameterized by its mean, $\mu_i$, and the Kullback-Leibler divergence between distributions $\nu_i$ and $\nu_i'$ can be written as $\KL(\nu_i, \nu_i') = \KL(\mu_i, \mu_i')$. In the fixed budget setting, any consistent algorithm satisfies:

$$
\limsup_{t \rightarrow \infty} -\frac{1}{t} \log p_t(\nu) \leq \max_{a \in \{1,...,K\}} \KL(\tau || \mu_a)
$$

or

$$
- \frac{1}{t} \log p_t(\nu) \leq \frac{1}{K}\sum_{i=1}^{K} \KL(\tau || \mu_i),
$$
\end{corollary}\label{corollary:LBKTBP}

We present the proofs of \autoref{sec:LBTBP} in the appendix in \autoref{sec:corollaryLBTBPproof}. Corresponding to the complexity measures defined in equation $(1)$ of Kaufmann et al. (2016), we can define the complexity $\kappa_{TBP}(\nu)$ in the thresholding bandit problem and lower bound it:
